Writing a reproducible paper with RStudio and Quarto

### 2023-03-08 paper.qmd
display tables only conditional for pdf or html.

`r if (knitr::is_latex_output()) "## Tables with stargazer()" else ""`

`r if (knitr::is_latex_output()) "@tbl-2 was generated using the stargazer() R package [@hlavac2013stargazer] and shows summary statistics of our data." else ""`

```{r, eval = knitr::is_latex_output()}
#| label: tbl-2
#| tbl-cap: "Summary table with stargazer"
#| results: "asis"
library(stargazer)
stargazer(swiss, 
          title = "Summary table with stargazer",
          label="tbl-2", 
          table.placement = "H", 
          type="latex",
          header=FALSE)
```

`r if (knitr::is_latex_output()) "In the above example we set the 'type' argument to 'latex' (default) which enables us to compile the table for PDF. This way, if you're reading the HTML output of this document, you won't be able to see the table (or it's cross-reference). This means, that dependent on your desired output format you need to change the type argument. For example, use 'type=html' for HTML output formats.<br> Lastly, note, how we specified 'results: asis' into the code chunk options. This is important so that latex is properly printed to the PDF." else ""`

@@

@tbl-4 shows a summary table for numeric variables.

```{{r}}
#| label: tbl-4
#| tbl-cap: "Summary: Numeric variables"
library(modelsummary)
datasummary_skim(swiss, 
                 type="numeric", 
                 histogram=T, 
                 title = "Summary: Numeric variables")
```

```{r eval=F}
#| label: tbl-4
#| tbl-cap: "Summary: Numeric variables"
library(modelsummary)
datasummary_skim(swiss, 
                 type="numeric", 
                 histogram=T, 
                 output="default",
                 title = "Summary: Numeric variables")
```

@@

@@

```{r}
#| label: tbl-5
#| tbl-cap: "Summary: Categorical variables"
#| results: "asis"
# Create categorical variables
swiss$Education_cat <- cut(swiss$Education, 
                   breaks=c(-Inf, 6, 12, Inf), 
                   labels=c("low","middle","high"))
swiss$Infant.Mortality_cat <- cut(swiss$Infant.Mortality, 
                   breaks=c(-Inf, 18.15, 21.70, Inf), 
                   labels=c("low","middle","high"))

library(flextable)
library(modelsummary)
tab_cat <- datasummary_skim(swiss, 
                            type="categorical", 
                            output = 'flextable')

@@

@tbl-6 shows the output for a regression table. Make sure you name all your models and explicitly refer to model names (M1, M2 etc.) in the text.

```{r echo=TRUE, paged.print=FALSE, results="asis"}
#| label: tbl-6
#| tbl-cap: "Linear regression"
library(stargazer)
M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
stargazer(M1, M2, M3,
          title = "Regression table with stargazer",
          label="tab2", 
          table.placement = "H", 
          column.labels = c("M1", "M2", "M3"),
          model.numbers = FALSE,
          header=FALSE)
```

## Regression table with modelsummary

@tbl-7 shows the output for a regression table. Make sure you name all the models you estimate (even if its 50) and explicitly refer to model names (M1, M2 etc.) in the text.

```{r eval=F}
#| label: tbl-7
#| tbl-cap: "Linear regression"
library(modelsummary)
M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
models <- list("M1" = M1, "M2" =  M2, "M3" = M3)

>>>>

## Tables with stargazer()

`r if (knitr::is_latex_output()) "@tbl-2-pdf was generated using the stargazer() R package [@hlavac2013stargazer] and shows summary statistics of our data." else ""`

`r if (knitr::is_html_output()) "@tbl-2-html was generated using the stargazer() R package [@hlavac2013stargazer] and shows summary statistics of our data." else ""`

```{{r}}
#| label: tbl-2
#| tbl-cap: "Summary table with stargazer"
#| results: "asis"
library(stargazer)
stargazer(swiss, 
          table.placement = "H", 
          type="latex", # adjust to html for HTML output
          header=FALSE)
```

```{r, eval = knitr::is_latex_output()}
#| label: tbl-2-pdf
#| tbl-cap: "Summary table with stargazer"
#| results: "asis"
library(stargazer)
stargazer(swiss, 
          table.placement = "H", 
          type="latex",
          header=FALSE)
```

```{r eval = knitr::is_html_output()}
#| label: tbl-2-html
#| tbl-cap: "Summary table with stargazer"
#| results: "asis"
library(stargazer)
stargazer(swiss,
          table.placement = "H", 
          type="html",
          header=FALSE)
```

In the above example, note that the `type` argument defaults to `latex` and you have to insert `html` if you desire HTML output. Lastly, note, how we specified 'results: asis' into the code chunk options. This is important so that latex is properly printed to the PDF.

@@

`r if (knitr::is_latex_output()) "@tbl-4-pdf shows a summary table for numeric variables." else ""`

`r if (knitr::is_html_output()) "@tbl-4-html shows a summary table for numeric variables." else ""`

```{{r}}
#| label: tbl-4
#| tbl-cap: "Summary: Numeric variables"
library(modelsummary)
datasummary_skim(swiss, 
                 type="numeric", 
                 histogram=T,
                 output="flextable", # adjust to html for HTML output
                 title = "Summary: Numeric variables")
```

```{r eval = knitr::is_latex_output()}
#| label: tbl-4-pdf
#| tbl-cap: "Summary: Numeric variables"
library(modelsummary)
datasummary_skim(swiss, 
                 type="numeric", 
                 histogram=T, 
                 output="flextable",
                 title = "Summary: Numeric variables")
```

```{r eval = knitr::is_html_output()}
#| label: tbl-4-html
#| tbl-cap: "Summary: Numeric variables"
library(modelsummary)
datasummary_skim(swiss, 
                 type="numeric", 
                 histogram=T, 
                 output="html",
                 title = "Summary: Numeric variables")
```

@@

library(flextable)
tab_cat <- datasummary_skim(swiss, 
                            type="categorical", 
                            output = 'flextable')  # adjust to html for HTML output

@@

```{r}
#| label: tbl-5
#| tbl-cap: "Summary: Categorical variables"

# Create categorical variables
swiss$Education_cat <- cut(swiss$Education, 
                   breaks=c(-Inf, 6, 12, Inf), 
                   labels=c("low","middle","high"))
swiss$Infant.Mortality_cat <- cut(swiss$Infant.Mortality, 
                   breaks=c(-Inf, 18.15, 21.70, Inf), 
                   labels=c("low","middle","high"))

#additionally we want to change the font, fontsize and spacing
library("gdtools")

@@

`r if (knitr::is_latex_output()) "@tbl-6-pdf shows a summary table for numeric variables." else ""`

`r if (knitr::is_html_output()) "@tbl-6-html shows a summary table for numeric variables." else ""`

```{{r}}
#| label: tbl-6
#| tbl-cap: "Linear regression with stargazer"
#| results: "asis"
library(stargazer)
M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
stargazer(M1, M2, M3,
          type="latex", # adjust to html for HTML output
          table.placement = "H", 
          column.labels = c("M1", "M2", "M3"),
          model.numbers = FALSE,
          header=FALSE)
```

```{r eval = knitr::is_latex_output()}
#| label: tbl-6-pdf
#| tbl-cap: "Linear regression with stargazer"
#| results: "asis"
library(stargazer)
M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
stargazer(M1, M2, M3,
          type="latex", # adjust to html for HTML output
          table.placement = "H", 
          column.labels = c("M1", "M2", "M3"),
          model.numbers = FALSE,
          header=FALSE)
```

```{r eval = knitr::is_html_output()}
#| label: tbl-6-html
#| tbl-cap: "Linear regression with stargazer"
#| results: "asis"
library(stargazer)
M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
stargazer(M1, M2, M3,
          type="html",
          table.placement = "H", 
          column.labels = c("M1", "M2", "M3"),
          model.numbers = FALSE,
          header=FALSE)
```

`r if (knitr::is_html_output()) "## Regression table with modelsummary" else ""`

`r if (knitr::is_html_output()) "@tbl-7 (HTML only) shows the output for a regression table. Make sure you name all the models you estimate (even if its 50) and explicitly refer to model names (M1, M2 etc.) in the text." else ""`

```{{r}}
#| label: tbl-7
#| tbl-cap: "Linear regression with modelsummary"
library(modelsummary)
M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
models <- list("M1" = M1, "M2" =  M2, "M3" = M3)

```{r eval = knitr::is_html_output()}
#| label: tbl-7
#| tbl-cap: "Linear regression with modelsummary"
library(modelsummary)
M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
models <- list("M1" = M1, "M2" =  M2, "M3" = M3)
